---
title: "First Quarto Report"
subtitle: "some Subtitle"
date: last-modified # display the date of the last edit
author: 
  - name: "Seraina"
    #orcid: 1234-1234-1234-12345
    affiliation: "University of Zurich"
format:
  html:
    toc: true # add a table of contents
  pdf: #default
    warning: false
    message: false
#execute:
 # warning: false
bibliography: contents/module_2_dynamic_reporting/demo_R/references.bib
---

## Quarto

Quarto enablesabout Quarto see <https://quarto.org>.

# This is a title

Two stars will render a word **bold**.

We can also create lists:

-   Element 1
-   Element 2

A useful link is this [Markdown guide](https://quarto.org/docs/authoring/markdown-basics.html)

> An impressive quote

```{r}
library("dplyr") # data wrangling
library("ggplot2") # plotting
library("kableExtra") # tables
```

```{r}
url <- "https://data.stadt-zuerich.ch/dataset/bev_vornamen_baby_od3700/download/BEV370OD3700.csv"
dat <- read.csv(url)
datclean <- dat |>
    rename(year = StichtagDatJahr,
           name = Vorname,
           sex = SexLang,
           births = AnzGebuWir) |>
    mutate(sex = ifelse(sex == "weiblich", "female", "male"))
```

```{r summarize data}
summaries <- datclean |>
  group_by(sex, year) |>
  slice_max(order_by = births, n = 1) |>
  arrange(-year)

summaries |>
  kable(caption = "Most common names of newborns in the city of Zurich by year and sex")
```
...

```{R}
datclean |>
  filter(year > 2020) |>
  group_by(sex, name) |>
  summarise(births = sum(births)) |>
  slice_max(order_by = births, n = 10) |>
  ggplot(aes(x = reorder(name, births), y = births)) +
  facet_wrap(~ sex, scales = "free") +
  geom_bar(stat = "identity") +
  labs(x = "Name", y = "Births", 
       title = "") +
  coord_flip() +
  theme_minimal()
```

Plot the frequency of names over time.

Name frequency is shown in @fig-name_graph.
```{R fig-name_graph}
datclean |>
  group_by(sex, name, year) |>
  filter(name %in% c("Fabio", "Daniel", "Seraina")) |>
  ggplot(aes(x = year, y = births, color = name)) +
  geom_line() +
  geom_point() +
  scale_color_viridis_d() +
  labs(x = "Year", y = "Births", color = "") +
  expand_limits(y = 0) +
  theme_minimal()
```


# Reproducibility

When creating dynamic reports, it is good practice to document the software versions and other parts of the computational environment. A simple yet effective way to document your computational environment is to include the output of sessionInfo().

...

```{r}
sessionInfo()
```

We can then use [@Hofmann2023] (with parentheses) or @Hofmann2023 (without parentheses) to cite a bibliography item named Hofmann2023



.

