---
title: Demo 
#listing:
 # categories: true
#  type: table
#  id: main-listing
#  fields: [title, subtitle]
#  sort-ui: true
#  filter-ui: true
  #contents: *.qmd
format:
  html:
    toc-depth: 4
  pdf: default
  #revealjs: default
execute: 
  eval: true
bibliography: references.bib
---

This demo requires R and Quarto Visit the [instructions](../instructions/index.qmd) page to get ready.

# Demo I - Create your own report

## 1. Install additional dependencies and packages

[Run the code below in the **R console** (it may take a couple of minutes)]{style="color: red;"}

``` r
 
# Install packages 
install.packages(c('quarto', 'dplyr', 'ggplot2','kableExtra'))
```

## 2. Start a new Quarto document

### Start a new QMD file

-   In the Rstudio menu, go to `File > New File > Quarto document`.
-   Keep the default options, fill in the filename text box
-   **Save** the file with a filename ending `.qmd` in the project home directory (visible on the bottom right pane).
-   Switch between the `visual` and `source` **tabs** of the R studio visual editor
-   Edit the code in the **source** tab.

### Edit the YAML header

Edit the YAML header delimited by (`---`). Add information about `author`, `affiliation` and `date` (use `last-modified` to always display the latest date) and define the output format as `pdf` or `html`. Beware of the *text indentation*! The header can be kept simple or it can contain a lot of information. An example:

``` yaml
---
title: 'Title of the report or page'
subtitle: 'Subtitle'
date: last-modified # Display the date of the last edit
author: 
  - name: 'Author 1 name'
    orcid: 1234-1234-1234-12345
    affiliations:
      - name: 'Institute, Faculty names'

  - name: 'Author 1 name'
    orcid: 1234-1234-1234-12345
    affiliations:
      - name: 'Institute, Faculty names'
format:
  html:
    toc: true
    toc-depth: 4
---
```

## 3. Create the report content

### Code chunks

The following [code chunks](https:/quarto.org/docs/computations/r.html) have executable code to perform different actions on the data

::: {#tip-chunks .callout-tip collapse="false"}
#### Code execution options

The code chunks are delimited by ```` ```{r} ```` (you can specify other languages like Python, Julia, ObservableJS). For each code you can add *execution options* to the code chunks with `#|`, for example `#| echo: false` will not display the code. In Rstudio you can see the options available if you type `#|` and then press the TAB ⭾ key on your keyboard . These options can be set globally for the entire document on the YAML header (under `execution:`).
:::

#### Load libraries

Let us first load all necessary libraries. [Copy the code below inside a 'code chunk']{style="color: red;"}

```{r}
#| warning: false
#| output: false

# Load libraries
lapply(c('dplyr', 'ggplot2','kableExtra'), library, character.only = TRUE)

```

#### Read the data

For this example we will use the data set 'Funding and Research Output Data' <https://zenodo.org/records/5011201>. The dataset is a table (CSV) with longitudinal data related to research outputs of researchers. Each researcher is identified with the variable `anonym_id` and there are multiple rows per researcher (*long format*).

We can read the table into a variable named *tbl* with the following code :

```{r}

url <- 'https://zenodo.org/records/5011201/files/anonymised_funding_output_data.csv?download=1'
filename <- 'anonymised_funding_output_data.csv'
tbl <- read.csv(url, row.names = 1)
```

#### Summary of the data

The data is longitudinal, we will summarize the cumulative sum of `n_articles_per_year` and `n_preprints_per_year` per researcher ( `anonym_id`).

```{r}

data_summary <- summarise(
                group_by(tbl, anonym_id),
                total_articles = sum(n_articles_per_year, na.rm = TRUE),
                total_preprints = sum(n_preprints_per_year, na.rm = TRUE)
           )
```

#### Table

We can render a table summarizing the data. For example:

```{r}
#| warning: true
#kableExtra::kable(head(tbl),format = "markdown")
kable(summary(data_summary)[,2:3], 
      caption = paste0('Summary of table with ', nrow(data_summary), ' rows'),
      booktabs =T) %>% kable_styling(latex_options=c("striped"))

```

#### Plot

We can add a code chunk a plot how the number of preprints developed over time. Note that we add some code execution options for a caption and a label that we can use to reference to this figure in the text.

```{r}
#| label: fig-preprints
#| fig-cap: "Number of preprints over the years."
#| echo: fenced


preprints <- summarise(
  group_by(tbl, year),
  n_preprints = sum(n_preprints_per_year)
  )

ggplot(preprints, aes(x= year, y = n_preprints)) + 
  geom_point(shape=21, fill = "blue") +
  theme_minimal()
```

I can now refer to the [@fig-preprints] in the text using `[@fig-label]`.

### Inline code

Now we can use this data_summary and [inline code](https://quarto.org/docs/computations/inline-code.html) to add a text describing the data using code to read some information from the table.

[Use code snippets such as those below in your Markdown text to describe the data]{style="color: red;"}

Total number of researchers: \``` r count(data_summary)` ``

Maximum number of publications: \``` r max(data_summary$total_articles)` ``

Median number of publications: \``` r median(data_summary$total_preprints)` ``

### References

You can refer to a file with the reference list in the YAML header. In this example with have the .bib file in the same folder as this .qmd file:

``` yaml
bibliography: references.bib
```

Then use `[@authorDate]` to add a citation. For example, we recommend reading this manifesto for reproducible science [@munafo_manifesto_2017] or this nice easy-read entitled `cut with the tyranny of copy-pasting` [@perkel_cut_2022] (yes, it is about R Markdown). 

To get the best out of this functionality 

-   [Connect your R studio project with a Zotero library](https://quarto.org/docs/visual-editor/technical.html#citations-from-zotero)

-   Use the Visual editor in R studio and go to insert>citation


## 4. Render the report

Now that you edit the QMD file with your report:

-   [ ] Check that the YAML header, the text and the code chunks are well delimited and you did not forget, for example, to close any chunk with (\`\`\`).
-   [ ] Click on the button `Render` with the blue arrow on top of the editor in R studio.
-   [ ] A file with the same name as your .qmd file and the format specified in your header should have been created.

## 5. Refine the report

Try to explore different options for code execution:

-   [ ] Hide all
-   [ ] Print all code but do not execute it
-   [ ] Fold all code (accessible by clicking on a button when saved an HTML)
-   [ ] Hide the code reading the table, show the rest
-   [ ] **Challenge** render the report as both an HTML document and as PDF

{{< pagebreak >}}

# Demo II - Some good habits in data visualisation

In the following we try illustrate some data visualisation principles by virtue of an example. TIt is based on the [CRS Primer on Data Visualisation](https://doi.org/10.5281/zenodo.8354336). The data is available here: <https://osf.io/rk3mp>. In addition we use ideas from @weissgerber_2015.

The primer identifies five principles of data visualisation:

-   Honesty\
-   Preattentive attributes\
-   Aesthetics\
-   Hierarchy\
-   Economy

We start from @fig-original_plot, comparing the mean outcomes in two groups with different treatment. We first load additionally required R libraries, then the data.

```{r}
#| warning: false
#| include: false
library(tidyr)
library(readr)
library(ggbeeswarm)
```

```{r}
#| label: load_data
#| include: false
vis_data <- read_csv("https://osf.io/download/rk3mp/")
```

```{r}
#| label: fig-original_plot
#| echo: true

data_summary <- vis_data %>% 
  group_by(NameOfDrug) %>% 
  summarise(value = mean(logOutc_value), sd = sd(logOutc_value)) 
  
p1 <- ggplot(data_summary, aes(x = NameOfDrug, y = value, fill = NameOfDrug)) +
  geom_bar(stat = "identity", alpha=0.7, width=0.5) +
  geom_pointrange(
    aes(ymin = value-sd, ymax = value+sd),
    color="gray40",
    linewidth = 1.3,
    size = 1.5
    ) +
  theme(
    panel.grid = element_line(color="lightgrey"),
    legend.title = element_text(size = 13),
    legend.text = element_text(size = 11)
  ) +
  labs(y = "Outcome", x = "Drug") +
  scale_fill_manual(values = c("red", "green"), name = "Drug") +
  coord_cartesian(ylim = c(7, 9.5)) +
  scale_y_continuous()

p1
```

## Improve the bar plot

We can improve this plot along some of the data visualisation principles introduced above.

-   Economy: 
    -   set a minimalist theme
    -   remove the redundant legend
    -   remove the panel grid
    -   remove axis.ticks
-   Preattentive attributes
    -   use more easily distinguishable colors
        -   A simulator for inspecting color palettes under different color vision conditions is available here: <https://www.susielu.com/data-viz/viz-palette>
        -   @zeileis_2020 provides a comprehensive discussion of using colors in R. 
-   Honesty:
    -   set axis to full range
    -   display aspects of the data distribution, not only summary statistics
        -   use boxplot instead of barplot
    -   **display raw data**
-   Aesthetics:
    -   use beeswarm or jitter
    -   color only the shapes' borders (use the aesthetic 'color' instead of 'fill')
    -   set appropriate axis font sizes
    
```{r}
#| label: improved_plot
#| eval: true
#| include: true
#| echo: true
p2 <- ggplot(vis_data, aes(x=NameOfDrug,y=logOutc_value,color=NameOfDrug))+
  geom_boxplot(width=0.1,position = position_nudge(x=0.3))+
  geom_beeswarm()+
  theme_bw()+
  theme(
    panel.grid.major = element_blank(),
    axis.text = element_text(size=14),
    axis.title = element_text(size=17),
    legend.position = "none"
  )+
  scale_color_manual(values=c("#00608b","#ffa600"),name="Drug")+
  labs(y="Outcome",x="Drug")+
  scale_y_continuous(limits=c(7.6,10),breaks=seq(7.5,10,0.5))

p2
```    

## Putting the two graphs together

We can use Quarto functionality to display the original and the revised plot side-by-side, specifically the execution options `layout-ncol` to define the number of columns, `fig-cap` for the global caption, and `fig-subcap` for column-specific captions.

```{r}
#| label: fig-summary
#| echo: false
#| layout-ncol: 2 # or layout: [[40, -10, 40]]
#| fig-cap: "Two figures that display the same data."
#| fig-subcap:
#|   - "Original figure"
#|   - "New figure"

p1
p2

```

# Demo III - Explore other examples

The are many examples of dynamically generated reports online. Try exploring the links below.

-   Example 1. Reproducible report in HTML using R Markdown (SNSF report: <https://data.snf.ch/stories/open-research-data-2023-en.html>)

-   Example 2. Reproducible report in pdf using Latex (robustness check analysis: <https://gitlab.uzh.ch/fabio.molo/nhb-replication>.)

-   Example 3. Website with course materials: <https://github.com/avehtari/BDA_course_Aalto/tree/master>\

-   Example 4. Reproducible slides with Quarto: <https://github.com/kjhealy/flipbookr-quarto/tree/main?tab=readme-ov-file>.\

Try to reflect about the following questions:

-   [ ] Can you easily find the repository with the source files?
-   [ ] What dynamic reporting tool they use?
-   [ ] Can you access the output, rendered documents easily ?
-   [ ] Do they use some advanced features for dynamic reporting or computational reproducibility?
-   [ ] What would be the workflow to reproduce those reports?

### references
