---
title: Demo
format:
  html:
    toc-depth: 4
  pdf: default
  #revealjs: default
execute: 
  eval: true
bibliography: references.bib
---

This demo requires R and Quarto Visit the [instructions](../instructions/index.qmd) page to get ready.

# Demo I - Create your own report

## 1. Install additional dependencies and packages

[Run the code below in the **R console** (it may take a couple of minutes)]{style="color: red;"}

``` r
install.packages(c("dplyr", "ggplot2", "kableExtra"))
```

## 2. Create a new Quarto document

### Start a new QMD file

-   In the Rstudio menu, go to `File > New File > Quarto document`.
-   Keep the default options, fill in the filename text box
-   **Save** the file with a filename ending `.qmd` in the project home directory (visible on the bottom right pane).
-   Switch between the `visual` and `source` **tabs** of the R studio visual editor
-   Edit the code in the **source** tab.

### Edit the YAML header

Edit the YAML header delimited by (`---`). Add information about `author`, `affiliation` and `date` (use `last-modified` to always display the latest date) and define the output format as `pdf` or `html`. Beware of the *text indentation*! The header can be kept simple or it can contain a lot of information. An example:

``` yaml
---
title: "Title of the report or page"
subtitle: "Subtitle"
date: last-modified # Display the date of the last edit
author: 
  - name: "Author 1 name"
    orcid: 1234-1234-1234-12345
    affiliations:
      - name: "Institute, Faculty names"

  - name: "Author 1 name"
    orcid: 1234-1234-1234-12345
    affiliations:
      - name: "Institute, Faculty names"
format:
  html:
    toc: true
    toc-depth: 4
---
```

## 3. Create the report content

### Text chunks

Quarto uses [Markdown syntax](https://quarto.org/docs/authoring/markdown-basics.html) to write and format text.

### Code chunks

The following [code chunks](https:/quarto.org/docs/computations/r.html) have executable code to perform different actions on the data

::: {#tip-chunks .callout-tip collapse="false"}
#### Code execution options

The code chunks are delimited by ```` ```{r} ```` (you can specify other languages like Python, Julia, ObservableJS). For each code you can add *execution options* to the code chunks with `#|`, for example `#| echo: false` will not display the code. In Rstudio you can see the options available if you type `#|` and then press the TAB ⭾ key on your keyboard . These options can be set globally for the entire document on the YAML header (under `execute:`).
:::

#### Load libraries

Let us first load all necessary libraries. [Copy the code below inside a "code chunk"]{style="color: red;"}

```{r}
#| warning: false
#| output: false
library("dplyr") # data wrangling
library("ggplot2") # plotting
library("kableExtra") # tables
```

#### Read the data

For this example we will use the data set "Vornamen neugeborener Mädchen und Knaben mit Wohnsitz in der Stadt Zürich" (<https://data.stadt-zuerich.ch/dataset/bev_vornamen_baby_od3700>). This data set contains the names of newborn children in the city of Zurich since the year 1993. The data was kindly made available to the public domain via a CC Zero license by Open Data Zürich.

We will now import the data and rename the variables into English.

```{r}
url <- "https://data.stadt-zuerich.ch/dataset/bev_vornamen_baby_od3700/download/BEV370OD3700.csv"
dat <- read.csv(url)
datclean <- dat |>
    rename(year = StichtagDatJahr,
           name = Vorname,
           sex = SexLang,
           births = AnzGebuWir) |>
    mutate(sex = ifelse(sex == "weiblich", "female", "male"))
```

#### Summaries the data

We will now compute the most common name by year and sex.

```{r}
summaries <- datclean |> 
  group_by(sex, year) |>
  top_n(n = 1, wt = births) |>
  arrange(year)
```

#### Create a table

We can render a table summarizing our results.

```{r}
summaries |>
  kable(caption = "Most common name of newborns in the city of Zurich by year and sex")
```

#### Plot the data

We can add a plot on the most common names between 2021 and 2023. Note that we add some code execution options for a caption and a label that we can use to cite this figure in the text.

```{r}
#| label: fig-names
#| fig-cap: "Most common names of Zurich newborns in 2021-2023."
#| message: false
datclean |> 
  filter(year > 2020) |>
  group_by(sex, name) |>
  summarise(births = sum(births)) |>
  top_n(n = 10, wt = births) |>
  ggplot(aes(x = reorder(name, births), y = births)) +
  facet_wrap(~ sex, scales = "free") +
  geom_bar(stat = "identity") +
  labs(x = "Name", y = "Births", 
       title = "") +
  coord_flip() +
  theme_minimal()
```

I can now refer to [@fig-names] in the text using `[@fig-label]`.

### Inline code

Now we can use this data_summary and [inline code](https://quarto.org/docs/computations/inline-code.html) to add a text describing the data using code to read some information from the table.

[Use code snippets such as those below in your Markdown text to describe the data]{style="color: red;"}

> In 2023, the most common male name was \``` r filter(summaries, year == 2023, sex == "male")$name` `` whereas the most common female name was \``` r filter(summaries, year == 2023, sex == "female")$name` ``.

produces

> In 2023, the most common male name was `r filter(summaries, year == 2023, sex == "male")$name` whereas the most common female name was `r filter(summaries, year == 2023, sex == "female")$name`.

### References

You can refer to a file with the reference list in the YAML header. In this example with have the .bib file in the same folder as this .qmd file:

``` yaml
bibliography: references.bib
```

Then use `[@authorDate]` to add a citation. For example, we recommend reading this manifesto for reproducible science [@munafo_manifesto_2017] or this nice easy-read entitled "cut with the tyranny of copy-pasting" [@perkel_cut_2022] (yes, it is about R Markdown).

To get the best out of this functionality

-   [Connect your R studio project with a Zotero library](https://quarto.org/docs/visual-editor/technical.html#citations-from-zotero)

-   Use the Visual editor in R studio and go to insert\>citation

### Reproducibility

It is good practice to save details on software versions and computational environment when creating dynamic reports. A basic but effective step to achieve this is the `sessionInfo` command.

```{r}
#| fontsize: "small"
sessionInfo()
```

## 4. Render the report

Now that you edit the QMD file with your report:

-   [ ] Check that the YAML header, the text and the code chunks are well delimited and you did not forget, for example, to close any chunk with (\`\`\`).
-   [ ] Click on the button `Render` with the blue arrow on top of the editor in R studio.
-   [ ] A file with the same name as your .qmd file and the format specified in your header should have been created.

## 5. Refine the report

Try to explore different options for code execution:

-   [ ] Hide all
-   [ ] Print all code but do not execute it
-   [ ] Fold all code (accessible by clicking on a button when saved an HTML)
-   [ ] Hide the code reading the table, show the rest
-   [ ] **Challenge** render the report as both an HTML document and as PDF

# Demo II - Explore other examples

The are many examples of dynamically generated reports online. Try exploring the links below.

-   Example 1. Reproducible report in HTML using R Markdown (SNSF report: <https://data.snf.ch/stories/open-research-data-2023-en.html>)

-   Example 2. Reproducible report in pdf using knitr (robustness check analysis: <https://gitlab.uzh.ch/fabio.molo/nhb-replication>.)

-   Example 3. Website with course materials: <https://github.com/avehtari/BDA_course_Aalto/tree/master>

-   Example 4. Reproducible slides with Quarto: <https://github.com/kjhealy/flipbookr-quarto/tree/main?tab=readme-ov-file>.

Try to reflect about the following questions:

-   [ ] Can you easily find the repository with the source files?
-   [ ] What dynamic reporting tool they use?
-   [ ] Can you access the output, rendered documents easily ?
-   [ ] Do they use some advanced features for dynamic reporting or computational reproducibility?
-   [ ] What would be the workflow to reproduce those reports?